---
title: "Visual Analysis of US Broadband Disparities"
output: 
  html_document:
    code_folding: hide
    includes: 
      before_body: header.html
    css: style.css
references:
  - id: fingas
    title: Microsoft Says the Rural Broadband Divide Is Worse than You Think
    author: 
    - family: Fingas
      given: Jon
    container-title: Engadget
    URL: "https://www.engadget.com/2018/12/05/microsoft-study-on-rural-broadband-shortfall/"
    type: article
    issued:
      year: 2018
  
  - id: kang
    title: How to Give Rural America Broadband? Look to the Early 1900s
    author: 
    - family: Kang
      given: Cecilia
    container-title: The New York Times
    URL: "https://www.nytimes.com/2016/08/08/technology/how-to-give-rural-america-broadband-look-to-the-early-1900s.html"
    type: article
    issued:
      year: 2017
      
  - id: lenz
    title: "Iowa: Rural Broadband, and the Unknown Costs of the Digital Divide"
    author: 
    - family: Lenz
      given: Lyz
    container-title: Colombia Journalism Review
    URL: "https://www.cjr.org/special_report/midterms-2018-iowa-rural-broadband.php/"
    type: article
    issued:
      year: 2018
---

## Background

As online services and devices become increasingly ingrained and necessary in our lives in the modern world, the Internet infrastructure and service providers that connect our devices to these online services has become increasingly important as well. Americans depend on reliable and fast broadband internet to conduct their business everyday in the 21st century. 

Yet, despite this necessity, millions of Americans have little to no access to broadband internet, most of them in rural areas where Internet Service Providers deem it unprofitable to provide broadband service [@fingas]. A lack of fast broadband Internet also means a lack of new business investment in the community, as most modern businesses require fast, reliable internet just to run, causing many rural communities to take things into their own hands [@kang]. 

These issues are magnified by the gross inaccuracies often found on reports analyzing broadband connectivity, creating little motivation by policy makers to fix things [@lenz]. This, and other factors like unaffordable broadband prices, exacerbates inequality, by restricting opportunities for communication, education, and employment to those who need them most, like unemployed citizens [@fingas]. 

## Data
| Name/Link                          | Rows      | Vars  | Unit          |
|------------------------------------|-----------|-------|---------------|
| [NYS Broadband Availability](https://data.ny.gov/Economic-Development/Broadband-Availability-By-Municipality/sjc6-ftj4)         | 1,635       | 24    | Municipality  |
| [People Without Internet](https://www.kaggle.com/madaha/people-without-internet)            | 821        | 23    | Municipality  |
| [Urban Rate Broadband Survey](https://public.opendatasoft.com/explore/dataset/urban-rate-broadband-survey/table/)        | 9,122     | 14    | ISP in Market |
| [American Community Survey 2017 ](https://www.census.gov/programs-surveys/acs/)        | 3,108     | 36    | County |
```{r loading-packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(ggplot2)
library(here)
library(ggmap)
library(maps)
library(mapdata)
library(stringr)
library(scales)
library(ggnewscale)
library(colorspace)
library(multiscales)
library(sf)
library(hexbin)
library(RColorBrewer)
library(reshape2)
library(readxl)
library(treemapify)

source("theme.R")

windows.options(antialias = "cleartype")
options(device = Cairo::CairoWin)

# install.packages("colorspace", repos = "http://R-Forge.R-project.org")
devtools::install_github("clauswilke/multiscales")
knitr::opts_chunk$set(fig.width=10, fig.height=7) 
```

##People Without Internet (American Community Survey 2016)
```{r acs-16-1, include=FALSE, cache=TRUE}
people_net = read_csv(here("Datasets", "US_People_Without_Internet.csv"))
state_regions = bind_rows(filter(state.fips, !str_detect(polyname, ":")), 
                          filter(state.fips, str_detect(polyname, ":main")))[4:5] %>% 
  bind_rows(tibble(division = c(9, 9), abb = c("AK", "HI")))
state_regions$division = state_regions$division  %>% factor()
levels(state_regions$division) = list(Northeast = c(1,2), Midwest = c(3, 4), South = c(5, 6, 7), West = c(8, 9))

people_net = merge(people_net, tibble(region = state_regions$division, state = state_regions$abb))

people_net$percent_below_poverty = (people_net$P_below_poverty / people_net$P_total) * 100
cor(people_net$percent_no_internet, people_net$percent_below_poverty)
people_net$percent_white = (people_net$P_white / people_net$P_total) * 100
cor(people_net$percent_no_internet, people_net$percent_white, use = "complete.obs")
```


```{r acs-16-2, cache=TRUE}
ggplot(data = people_net, aes(x = percent_no_internet, y = percent_below_poverty)) + 
  geom_point(aes(size = P_total, color = region), alpha = 0.6) + 
  scale_color_manual(values = color_pal(4)) +
   scale_size(range = c(2, 8.5), labels = comma) +
  geom_smooth(method='lm',formula=y~x, color = "dark grey") + 
  labs(title = "Communities with Higher Poverty Rates Have Less Internet Access", 
       subtitle = "US Counties that have a higher percentage of people with no Internet also \nhave a higher percentage of residents below the poverty line", 
       x = "Percent with no Internet", 
       y = "Percent Below Poverty Line", 
       caption="American Community Survey 2016", 
       color = "Region", size = "Total Population") + 
  scale_x_log10(breaks = c(0, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50)) + 
  scale_y_log10(breaks = c(0, 5, 10, 15, 20, 25, 30)) + 
  theme_master()
```

As stated previously, there is evidence that a lack of Internet access can exacerbate inequality. While there is not enough evidence to make a causal relationship, the strong correlation between poverty and Internet access at least shows that a relationship does exist between the two. The lack of a clear trend between region or county population and this internet access - poverty relationship suggests that this relationship holds true across all of America.

```{r acs-16-3, cache=TRUE}
ggplot(data = people_net, aes(x = percent_no_internet, y = percent_below_poverty)) + 
  geom_smooth(method='lm',formula=y~x, color = "dark grey") + 
  geom_point(aes(size = P_total, color = region), alpha = 0.6) + 
  scale_color_manual(values = color_pal(4)) +
  scale_size(range = c(1.5, 7.5), labels = comma) +
  facet_wrap(~ region) +
  labs(title = "The West and South Have Larger Disparities in Inequality", 
       subtitle = "The Southern and Western US have more areas with better internet and less poverty and vice versa \nwhen compared to the Northeast and Midwest", 
       x = "Percent with no Internet", 
       y = "Percent Below Poverty Line", 
       caption="American Community Survey 2016", 
       color = "Region", size = "Total Population") + 
  scale_x_log10(breaks = c(0, 5, 10, 20, 35, 50)) + 
  scale_y_log10(breaks = c(0, 5, 10, 30)) + 
  theme_master()
```

However, despite the increased variance in the South and West, the relationship between Internet Access and Poverty holds true across the US. Perhaps the wider range of population density in the South and West exacerbates the disparity of this relationship. The extreme outlier in the lower left corner in Western America, is Douglas County, which is in the top 10 most populated counties in Colorado.

## Urban Broadband Survey

```{r ubs-setup, include=FALSE, cache=TRUE}
urban_bb = read_csv2(here("Datasets", "US_Urban_Rate_Broadband_Survey.csv"))
urban_bb$`Total Charge` = round(as.numeric(urban_bb$`Total Charge`))
urban_bb$`Download Bandwidth Mbps` = round(as.numeric(urban_bb$`Download Bandwidth Mbps`))
urban_bb$isp_size = -log(as.numeric(urban_bb$Weight))
urban_bb$isp_size = (urban_bb$isp_size - min(urban_bb$isp_size) + 1)^3


perc_not_bb = sum(urban_bb$`Download Bandwidth Mbps` < 25) / length(urban_bb$`Download Bandwidth Mbps`) * 100

urban_bb$Provider = recode(urban_bb$Provider,
                           `Charter Communications, Inc.` = "Charter", 
                             `Charter Communications` = "Charter", 
                             `COMCAST CABLE COMMUNICATIONS, INC.` = "Comcast",
                             `COMCAST CABLE COMMUNICATIONS, LLC` = "Comcast",
                             `Cox Communications` = "Cox",
                             `Cox Communications, Inc` = "Cox",
                             `DigitalPath, Inc` = "DigitalPath",
                             `DigitalPath, Inc.` = "DigitalPath",
                             `Frontier Communications` = "Frontier",
                             `Frontier Communications Corporation` = "Frontier",
                             `Giggle Fiber LLC` = "Google Fiber",
                             `Google Fiber Kansas, LLC` = "Google Fiber",
                             `Google Fiber Missouri, LLC` = "Google Fiber",
                             `Google Fiber Texas, LLC` = "Google Fiber",
                             `Google Fiber Utah, LLC` = "Google Fiber",
                             `Knology of Alabama, Inc.` = "Knology",
                             `Knology of Florida, Inc.` = "Knology",
                             `Liberty Cablevision of Puerto Rico LLC` = "Liberty Cablevision",
                             `Liberty Cablevision of Puerto Rico, LLC` = "Liberty Cablevision",
                             `Mediacom Arizona LLC` = "Mediacom",
                             `Mediacom California LLC` = "Mediacom",
                             `Mediacom Delaware LLC` = "Mediacom",
                             `Mediacom Illinois LLC` = "Mediacom",
                             `Mediacom Indiana LLC` = "Mediacom",
                             `Mediacom Iowa LLC` = "Mediacom",
                             `Mediacom Minnesota LLC` = "Mediacom",
                             `Mediacom Southeast LLC` = "Mediacom",
                             `Orlando Telephone Company, Inc.` = "Orlando Telephone Company",
                             `RCN Telecom Services (Lehigh) LLC` = "RCN",
                             `RCN Telecom Services of Illinois, LLC` = "RCN",
                             `RCN Telecom Services of New York, LP` = "RCN",
                             `RCN Telecom Services of Philadelphia LLC` = "RCN",
                             `RCN Telecom Services of Philadelphia, LLC` = "RCN",
                             `Time Warner Cable Enterprises LLC` = "Time Warner",
                             `Time Warner Cable Inc.` = "Time Warner",
                             `Verizon California Inc.` = "Verizon",
                             `Verizon Delaware LLC` = "Verizon",
                             `Verizon Florida LLC` = "Verizon",
                             `Verizon Maryland LLC` = "Verizon",
                             `Verizon New England Inc.` = "Verizon",
                             `Verizon New Jersey Inc.` = "Verizon",
                             `Verizon New York Inc.` = "Verizon",
                             `Verizon Pennsylvania Inc.` = "Verizon",
                             `Verizon Pennsylvania LLC` = "Verizon",
                             `Verizon Virginia Inc.` = "Verizon",
                             `Verizon Virginia LLC` = "Verizon",
                             `Verizon Washington, DC Inc.` = "Verizon",
                             `Windstream Concord Telephone, Inc` = "Windstream",
                             `Windstream Corporation` = "Windstream",
                             `Windstream D&E, Inc.` = "Windstream",
                             `Windstream Georgia Communications, LLC` = "Windstream",
                             `Windstream Kentucky East, LLC` = "Windstream",
                             `Windstream Mississippi, LLC` = "Windstream",
                             `Windstream Missouri, Inc.` = "Windstream",
                             `Windstream Nebraska, Inc.` = "Windstream",
                             `Windstream North Carolina, LLC` = "Windstream",
                             `Windstream Ohio, Inc.` = "Windstream",
                             `Windstream Pennsylvania, LLC` = "Windstream",
                             `Windstream South Carolina, LLC` = "Windstream",
                             `Windstream Sugar Land, Inc` = "Windstream",
                             `Windstream Western Reserve, Inc` = "Windstream",
                           `AT&T Services, Inc.` = "AT&T", 
                           `CenturyLink, Inc.` = "CenturyLink"
         )
urban_bb$down_value = urban_bb$`Download Bandwidth Mbps` / urban_bb$`Total Charge`

agg_urban_bb = NULL

for (type in unique(urban_bb$Technology)) {
  agg_func = function(x) {
    return(mean(as.numeric(x)))
  }
  
  temp = subset(urban_bb, Technology == type) %>% subset(`Total Charge` > 0)
  result = temp %>% .[c(4,6,7,8,9,10,11,12,13,17,18)] %>% 
  aggregate(by = list(temp$Provider), FUN = agg_func) %>% 
    bind_cols(tibble(counts = table(temp$Provider)))
  result$Technology = type
  if (is.null(agg_urban_bb)) {
    agg_urban_bb = result
  } else {
    agg_urban_bb = bind_rows(agg_urban_bb, result)
  }
}

agg_urban_bb = subset(agg_urban_bb, Technology != "Other")

```

<div class="dark_theme">
```{r ubs-1, message=FALSE, warning=FALSE, cache=TRUE}
chart_labels = tibble(text = c('FCC \nBroadband \nCutoff', 'Trendline'), x = c(8.5, 750), y = c(25, 2500))

ggplot(data = urban_bb, aes(x = `Total Charge`, y = `Download Bandwidth Mbps`)) + 
  geom_smooth(method='lm', formula=y~x, color = "grey") + 
  geom_point(color = color_pal(1), alpha = 0.5, size = 2.5) + 
  geom_hline(yintercept = 25, linetype = 2, color = "white", size = 1) + 
  geom_label(data = chart_labels, aes(x = x, y = y, label = text), alpha = 0.95) + 
  scale_y_continuous(trans = "log10") +
  scale_x_continuous(trans = "log10") +
  labs(title = "Faster Download Speeds Mean Costlier Internet", 
       subtitle = paste("Even in urban areas, ", round(perc_not_bb, 0), "% of survey respondents from 2015 - 2018 had internet slower \nthan the 25 Mbps minimum download speed for broadband as defined by the FCC", sep = ""),
       x = "Total Monthly Internet Cost (USD)",
       y = "Download Speed (Mbps)", 
       caption = "Urban Rate Broadband Survey") +
  theme_master_dark()
```

Existing literature suggests that rural areas have disproportionately less broadband Internet access than urban areas. However, even in urban areas, Internet access can be extremely expensive, even for access that does not meet the FCC's minimum bandwidth requirements to be legally considered broadband internet. While, for the most part, people can pay more to get faster speeds, inflated pricing disproportionately affects lower income families, who may become stuck with slow internet as it is all that they can afford.
</div>

```{r ubs-2, message=FALSE, warning=FALSE, cache=TRUE}
ggplot(data = filter(urban_bb, Technology != "Other"), aes(x = `Total Charge`, y = `Download Bandwidth Mbps`)) + 
  geom_smooth(method='lm', formula=y~x, color = "grey") + 
  geom_point(aes(color = Technology), alpha = 0.5, size = 2) + 
  scale_color_manual(values = color_pal(4)) +
  geom_hline(yintercept = 25, linetype = 2) + 
  facet_wrap(~ Technology) +
  scale_y_continuous(trans = "log10") +
  scale_x_continuous(trans = "log10") +
  labs(title = "Download Speeds and Costs Vary by Network Technology",
       x = "Total Monthly Internet Cost (USD)",
       y = "Download Speed (Mbps)", 
       caption = "Urban Rate Broadband Survey") +
  theme_master() +
  hide_legend
```

Naturally, download speeds vary by the technology that Internet Service Providers use in their network. Fiber to the Home offers the highest speeds, followed by cable, however both have a similar price to speed ratio. DSL is typically much slower, but also less expensive, while Fixed Wireless is, overall, more expensive for the same speeds compared to other technologies.

```{r ubs-tmap}
ggplot(agg_urban_bb, aes(area = as.numeric(counts), fill = down_value, label = Group.1, subgroup = Technology)) + 
  geom_treemap() +
  geom_treemap_subgroup_border() +
  geom_treemap_subgroup_text(place = "centre", grow = T, alpha = 0.5, colour =
                             "black", fontface = "italic", min.size = 0) +
  geom_treemap_text(colour = "white", place = "centre",
                    grow = TRUE) +
  scale_fill_gradientn(trans = "log10", colors = color_pal(5, type = "continuous")) + 
  labs(title = "Cable and Fiber ISPs Provide Better Value for Internet Service",
       subtitle = "Small ISPs that use Fiber to the Home are the best value",
       fill = "Value \n(Megabits \nper USD)", 
       caption = "Urban Rate Broadband Survey") +
  theme_master()
```

Overall, Fiber to the Home provides the best value for consumers, with cable not far behind. DSL and Fixed Wireless have the worst value on average. While small FTTH ISPs, like Google Fiber, typically provide better value than larger fiber ISPs, this is not true for other technologies, where larger ISPs usually provide better value. 

```{r ubs-3, message=FALSE, warning=FALSE, cache=TRUE}
ggplot(data = urban_bb, aes(x = `Total Charge`, y = `Download Bandwidth Mbps`)) + 
  geom_smooth(method='lm', formula=y~x, color = "grey") + 
  geom_point(aes(color = ordered(Year)), alpha = 0.5, size = 2) + 
  scale_color_manual(values = color_pal(4, "continuous")) +
  geom_hline(yintercept = 25, linetype = 2) + 
  facet_wrap(~ Year) +
  scale_y_continuous(trans = "log10") +
  scale_x_continuous(trans = "log10") +
  labs(title = "Download Speeds and Costs Vary by Year",
       subtitle = "Even as faster speeds have become more common over time, the relationship \nbetween price and download speed has changed very little.",
       x = "Total Monthly Internet Cost (USD)",
       y = "Download Speed (Mbps)",
       color = "Year",
       caption = "Urban Rate Broadband Survey") +
  theme_master() +
  hide_legend
```

```{r ubs-4, message=FALSE, warning=FALSE, cache=TRUE}
ggplot(data = urban_bb, aes(x = factor(`Year`), y = `Total Charge`)) +
  geom_violin(aes(color = ordered(Year), fill = ordered(Year))) +
  scale_color_manual(values = color_pal(4, "continuous")) +
  scale_fill_manual(values = color_pal(4, "continuous")) +
  scale_y_continuous(trans = "log10", breaks = c(10, 30, 50, 100, 150, 300, 1000)) +
  labs(title = "Monthly Internet Costs Have Increased Slightly",
       subtitle = "Most people paid around $50 for Internet access every month in 2015-2018. \nHowever, less people are paying less than $50 and more between $50-100 in 2018 than 2015.",
       x = "Year",
       fill = "Year",
       color = "Year",
       y = "Total Monthly Internet Cost (USD)",
       caption = "Urban Rate Broadband Survey") +
  theme_master() +
  theme(legend.position="none")
```
<div class="dark_theme">
```{r ubs-5, message=FALSE, warning=FALSE, cache=TRUE}
chart_labels = tibble(text = c('FCC \nBroadband \nCutoff'), x = c(0.75), y = c(25))
ggplot(data = urban_bb, aes(x = factor(`Year`), y = `Download Bandwidth Mbps`)) +
  geom_violin(aes(color = ordered(Year), fill = ordered(Year))) +
  scale_color_manual(values = color_pal(4, "continuous")) +
  scale_fill_manual(values = color_pal(4, "continuous")) +
  scale_y_continuous(trans = "log10", breaks = c(1, 5, 20, 50, 100, 300, 1000, 10000)) +
  geom_hline(yintercept = 25, linetype = 2,  color = "white", size = 1) + 
  geom_label(data = chart_labels[1,], aes(x = x, y = y, label = text), alpha = 0.95) + 
  labs(title = "Download Speeds Have Increased Significantly",
       subtitle = "High-speed Internet, especially Gigabit Internet subscribtions have increased noticeably between 2015 and 2018 \nStill, many people remained on Internet plans with speeds below the FCC minimum for broadband \nat less than 25 Mbps in 2018",
       x = "Year",
       fill = "Year",
       color = "Year",
       y = "Download Speed (Mbps)",
       caption = "Urban Rate Broadband Survey") +
  theme_master_dark() +
  theme(legend.position="none")
```

Even though prices have slightly increased, there is a clear trend in people purchasing faster internet over time. This perhaps indicates that either people are upgrading their internet to faster speeds as it become available, or that faster internet has become more of a necessity, causing people to be more willing to pay for better internet. Regardless, more people have internet that meets the FCC minimum requirements for broadband in 2018 than 2015.
</div>

```{r ubs-6, message=FALSE, warning=FALSE, cache=TRUE}
get_perc = function(data_set, category, n_var) {
  len = length(data_set[[1]])
  for (i in 1:len) {
    data_set$percent[i] = eval(parse(text = paste("data_set$", n_var, "[i] / sum(subset(data_set,", 
                                                 category, "== data_set$", 
                                                 category, "[i])$", 
                                                 n_var, 
                                                 ")",
                                                 sep = "")))
    
  }
  return(data_set)
}

urban_bb[c(2,5)] %>% table() %>% as_tibble() %>% get_perc("Year", "n") %>%
  ggplot(aes(x = as.numeric(Year), color = Technology, y = as.numeric(percent))) + 
  scale_color_manual(values = color_pal(5)) +
  geom_line(size = 1.5) + geom_point(size = 2) +
  labs(title = "More People are Choosing Fiber to the Home and Fixed Wireless Internet",
       subtitle = "The amount of survey respondents who had Cable or DSL Internet \nhas decreased between 2015 and 2018",
       x = "Year", 
       y = "Percentage of Respondents",
       caption = "Urban Rate Broadband Survey") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5)) +
  theme_master() 
```

More people are choosing Fiber to the Home for their internet access in 2018 than previous years. This offers evidence that fiber is becoming more accessible to people, either due to increased investment in fiber infrastructure by ISPs, or from more competitive pricing by ISPs that use Fiber to the Home technology.

```{r ubs-7, message=FALSE, warning=FALSE, cache=TRUE}
urban_bb$usage_limit = (urban_bb$`Usage Allowance GB` != "Unlimited")
urban_bb[c(2,19)] %>% table() %>% as_tibble() %>% get_perc("Year", "n") %>%
  ggplot(aes(x = as.numeric(Year), color = usage_limit, y = percent)) + 
  scale_color_manual(values = color_pal(2, reverse = TRUE), labels = c("No", "Yes")) +
  geom_line(size = 1.5) + geom_point(size = 2) +
  labs(title = "Data Caps are Not Disappearing",
       subtitle = "The amount of survey respondents who had Internet Service with a monthly data cap \nhas neither grown or decreased substantially between 2015 and 2018",
       x = "Year", 
       color = "Has Data Cap",
       y = "Percentage of Respondents",
       caption = "Urban Rate Broadband Survey") +
  scale_y_continuous(labels = percent_format(), limits = c(0, 0.65)) +
  theme_master() 
```
Data Caps restrict the Internet service a customer pays for, only allowing a certain amount of data to be transferred per month. This not only disproportionately affects lower income families, it can also make it difficult or impossible to "cut the cord", and switch from TV to video streaming services. Data caps also make it more expensive for small businesses, who often may need to upgrade to "business" internet to avoid overage charges.

## New York State Broadband Availability
```{r nys-prep, include=FALSE}
ny_bb = read_csv(here("Datasets", "NY_Broadband_Availability.csv")) 

ny_map_data = st_as_sf(map("county", plot = FALSE, fill = TRUE)) %>% separate(ID, c("state", "subregion"), sep = ",") %>%
  subset(state %in% "new york") %>% .[c(2,3)]
ny_map_data$subregion = ny_map_data$subregion %>% recode("st lawrence" = "st. lawrence")
  
ny_map = ny_map_data %>% ggplot() + geom_sf(color = "grey90", fill = "white")

ny_county_bb = subset(ny_bb, ny_bb$`Municipality Type` == "County") %>% rename("NAME" = "Municipality Name") 
ny_county_bb$subregion = tolower(ny_county_bb$NAME)
ny_county_bb = ny_map_data %>% 
  full_join(ny_county_bb)
ny_county_bb$pop_dens = ny_county_bb$`2010 Muni Population` / ny_county_bb$`Muni Area (sq mi)`
```

```{r nys-1, message=FALSE, warning=FALSE}
ny_county_bb$`# Fiber Providers` = ny_county_bb$`# Fiber Providers` %>% ordered()
levels(ny_county_bb$`# Fiber Providers`) = list("1" = c(1), "2" = c(2), "3" = c(3), "4" = c(4), "5" = c(5), "6+" = c(6,7,8,9) )
ggplot(data = ny_county_bb, 
       aes(y = pop_dens, x = as.factor(ny_county_bb$`# Fiber Providers`))) + 
  geom_violin(aes(fill = as.factor(ny_county_bb$`# Fiber Providers`), 
                  color = as.factor(ny_county_bb$`# Fiber Providers`))) + 
  scale_fill_manual(values = color_pal(6)) +
  scale_color_manual(values = color_pal(6)) +
  scale_y_continuous(trans = "log10") + 
  labs(title = "Urban Areas Have More Fiber Internet Choices",
       subtitle = "The more dense a New York county is, the more likely it is to have several fiber providers",
       x = "Number of Fiber Providers per County", 
       y = "Population Density of County (ppl/sq mi)",
       caption = "New York State Broadband Availability By Municipality") + 
  theme_master() +
  theme(legend.position="none")
```

```{r nys-2}
ggplot(data = ny_county_bb, 
       aes(y = pop_dens, 
           x = as.factor(ny_county_bb$`# Cable Providers`))) + 
  geom_violin(aes(fill = as.factor(ny_county_bb$`# Cable Providers`), 
                  color = as.factor(ny_county_bb$`# Cable Providers`))) + 
  scale_fill_manual(values = color_pal(4)) +
  scale_color_manual(values = color_pal(4)) +
  scale_y_continuous(trans = "log10") + 
  labs(title = "Urban Areas Have More Cable Internet Choices",
       subtitle = "The more dense a New York county is, the more likely it is to have several cable providers",
       x = "Number of Cable Providers per County",
       y = "Population Density of County (ppl/sq mi)",
       caption = "New York State Broadband Availability By Municipality") + 
  theme_master() +
  theme(legend.position="none")
```

This data from New York offers some evidence that there are more ISPs the more population dense a county is. This suggests that ISPs are hesitant to enter a new market in rural areas, either because of the high barrier of entry, or due to existing ISPs in the market protecting their investment through tactics like exclusivity agreements with municipalities. 

```{r nys-3, echo=TRUE}
ny_county_bb = ny_county_bb [rev(order(ny_county_bb $pop_dens)),]

ny_map + geom_sf(data = ny_county_bb, 
                 aes(fill = `% Hse Units Cable` / 100,
                     color = `% Hse Units Cable` / 100,
                     alpha = pop_dens),
                 size = 0.7) + 
 scale_fill_gradientn(colors = color_pal(5, type = "continuous"), 
                        labels = percent_format(accuracy = 1)) +
  scale_color_gradientn(colors = color_pal(5, type = "continuous"), 
                        labels = percent_format(accuracy = 1)) +
  scale_alpha(trans = "log", range = c(0.1, 1), limits = c(1, 50000), breaks = c(5, 20, 400, 8000)) +
  labs(title = "Rural Counties have Less Cable Internet Coverage",
       subtitle = "Less populous counties proportionately have less households with cable internet",
       fill = "Percent with \nCable Internet",
       color = "Percent with \nCable Internet",
       alpha = "Population Density \n(ppl/sq mi)",
       caption = "New York State Broadband Availability By Municipality") +
    theme_map() +
  theme(axis.text = element_blank(),
        legend.position = "right")
```

```{r nys-4}
ny_map + geom_sf(data = ny_county_bb, 
                 aes(fill = `% Hse Units Fiber` / 100,
                     color = `% Hse Units Fiber` / 100,
                     alpha = pop_dens),
                 size = 0.7) + 
  scale_fill_gradientn(colors = color_pal(5, type = "continuous"), 
                        labels = percent_format(accuracy = 1)) +
  scale_color_gradientn(colors = color_pal(5, type = "continuous"), 
                        labels = percent_format(accuracy = 1)) +
  scale_alpha(trans = "log", range = c(0.1, 1), limits = c(1, 50000), breaks = c(5, 20, 400, 8000)) +
  labs(title = "Rural Counties have Less Fiber Internet Coverage",
       subtitle = "Less populous counties proportionately have less households with fiber internet",
       fill = "Percent with \nFiber Internet",
       color = "Percent with \nFiber Internet",
       alpha = "Population Density \n(ppl/sq mi)",
       caption = "New York State Broadband Availability By Municipality") +
    theme_map() +
  theme(axis.text = element_blank(), 
        legend.position = "right")
```

While overall, New York has good cable internet coverage, there still are significant disparities between rural and urban counties when it comes to cable Internet access, according to the state's own broadband survey. However, only a few counties in New York City and the metro area have high fiber Internet adoption rates. 

## US Census - American Community Survey 2017

```{r import-acs17-data, message=FALSE, warning=FALSE, include=FALSE}
states_sf = read_sf(here("Datasets", "States_Shapefile", "cb_2017_us_state_20m.shx")) %>% filter(as.numeric(STATEFP) %in% c(1:56)[-c(2,15)])

us_sf = read_sf(here("Datasets", "US_Shapefile", "tl_2018_us_county.shx")) %>% filter(as.numeric(STATEFP) %in% c(1:56)[-c(2,15)])
# Oglala Lakota County, SD was Shannon County before 2015, which needs to be corrected, as apparently this area database ("LND01.xls") has not been updated since the name change
land_area = read_excel(here("Datasets", "LND01.xls")) %>% select(STCOU, LND110210D)
land_area[land_area$STCOU == 46113,]$STCOU = 46102
acs_17 = land_area %>% rename(GEOID = STCOU, land_area = LND110210D) %>% right_join(readRDS(here("Datasets", "acs_17_vals.rds"))) %>% inner_join(as_tibble(us_sf)[,c(4,16,17)], by = "GEOID")

acs_17_mod = acs_17[1:length(acs_17[[1]]) ,c(1,3,21,22,23,31,34,35,36)] %>%
  bind_cols(data.frame(pop_dens = (acs_17$total_pop / acs_17[,2]))) %>% 
  bind_cols(as_tibble(acs_17[, -c(1,2,3,21,22,23,31,34,35,36)] / acs_17$total_pop)) %>% 
  rename("pop_dens" = land_area, "avg_fam_inc" = aggr_fam_inc) %>%
  separate(NAME, c("county", "state"), sep = ", ")

acs_17_mod$INTPTLON = -as.numeric(substr(acs_17_mod$INTPTLON, start = 2, stop = 11))
acs_17_mod$INTPTLAT = as.numeric(substr(acs_17_mod$INTPTLAT, start = 2, stop = 11))
acs_17_mod = acs_17_mod[-8] %>% st_as_sf(coords = c("INTPTLON", "INTPTLAT"), crs = 4326)

transformed_points = acs_17_mod$geometry %>% st_transform(crs = 102003)
acs_17_mod = cbind(acs_17_mod, st_coordinates(transformed_points))

acs_17_mod = acs_17_mod[rev(order(acs_17_mod$pop_dens)),]
```

```{r mapping-internet-data} 
states_sf %>% 
  ggplot() + 
  geom_sf(color = "grey90", fill = "white") +
  coord_sf(crs = st_crs(102003)) +
  geom_point(data = acs_17_mod, aes(x = X,
                     y = Y,
                     color = (1- no_inet),
                     size = pop_dens)) + 
  scale_color_gradientn(colors = color_pal(6, type = "cool"), labels = percent_format(accuracy = 1)) +
  scale_size(range = c(0.5, 15), breaks = c(50, 500, 5000, 50000)) +
  scale_alpha(range = c(1.0, 0.05)) +
  labs(title = "Most of America has Internet Access",
       subtitle = "Even in rural counties, the majority of households have internet access",
      color = "Percent with \nInternet",
       size = "Population Density \n(ppl/sq mi)",
      caption = "American Community Survey 2017") +
   theme_map() +
  theme(axis.text = element_blank())
```
<div class="dark_theme">
```{r mapping-broadband-data} 
states_sf %>% 
  ggplot() + 
  geom_sf(color = "grey30", fill = "#323232") +
  coord_sf(crs = st_crs(102003)) +
  geom_point(data = acs_17_mod, aes(x = X,
                     y = Y,
                     color = broadband_any,
                     size = pop_dens)) + 
  scale_color_gradientn(colors = color_pal(6, type = "warm", reverse = TRUE), labels = percent_format(accuracy = 1)) +
  scale_size(range = c(0.5, 15), breaks = c(50, 500, 5000, 50000)) +
  scale_alpha(range = c(1.0, 0.05)) +
  labs(title = "Rural Counties have Less Access to Broadband Internet",
       subtitle = "Even in urban areas, many households, while having Internet, do not have broadband Internet access",
      color = "Percent with \nBroadband",
       size = "Population Density \n(ppl/sq mi)",
      caption = "American Community Survey 2017") +
  theme_map(dark_theme = TRUE) +
  theme(axis.text = element_blank())
```
In rural counties in the United States, while most households have Internet access, many do not have Internet that meets the minimum speed requirements to be considered broadband Internet. However, this varies across the country. For example, Utah has much higher percentages of households for overall Internet access and broadband Internet access than many other rural states. In fact, despite its vastness, the western United States overall has much better accessibility to high-speed internet than other areas, most notably the rural South. 

</div>

## Works Cited